### 2015-08-21 templates/report/research/tex/academic_works.tex
Testing tex file

{% block content %}

{% if postdoc_concluded or postdoc_in_progress or phd_concluded or phd_in_progress or msc_concluded or msc_in_progress %}
	{% if postdoc_concluded or postdoc_in_progress %}
\subsubsection{Post-doctoral fellows}
		{% if postdoc_concluded %}
\paragraph{Concluded}
\begin{enumerate}
			{% for postdoc in postdoc_concluded %}
	\item Title of the project: {{ postdoc.title }}
	\begin{itemize}
		\item Period: {{ postdoc.start_date }} to {{ postdoc.end_date }}
		\item Name of the student: {{ postdoc.advisee }}
		\item Advisor: {{ postdoc.advisor }}
		{% if postdoc.co_advisors %}\item Co-advisor: {{ postdoc.co_advisors }}{% endif %}
		\item Institution: {{ postdoc.institution }}
		{% if postdoc.schollarship %}\item Schollarship: {{ postdoc.schollarship }}{% endif %}
	\end{itemize}
             {% endfor %}
\end{enumerate}
		{% endif %}	
	{% endif %}
{%  else %}
    No academic works found!
{% endif %}

{% endblock %}