### 2016-04-01 templates/report/research/tex/articles.tex
NIRA-60 - Create an article report in Latex format

{% block content %}

{% if published_scientific or accepted_scientific or submitted_scientific or draft_scientific or published_dissemin or accepted_dissemin or submitted_dissemin or draft_dissemin or published_tec_trans or accepted_tec_trans or submitted_tec_trans or draft_tec_trans or published_scientific_in_event or accepted_scientific_in_event or published_dissemin_in_event or accepted_dissemin_in_event or published_tec_trans_in_event or accepted_tec_trans_in_event%}

	{% if published_scientific or accepted_scientific %}
\subsubsection{Published or accepted}
		{% if published_scientific %}
\begin{enumerate}
			{% for article in published_scientific %}
				{% if not article.article.hide %} \item {% if article.article.url %} \textbf{ \href{ {{ article.article.url }} }{ {{ article.article.title }}. } }{% else %}{{ article.article.title }}.{% endif %}{% if article.article.ris_file_authors  %} \textit{ {{ article.article.ris_file_authors }} }{{ article.article.periodical }},{% else %} {{ article.article.authors }} {{ article.article.periodical }},{% endif %}{% if article.volume %} \textbf{ {{ article.volume }} }{% endif %}{% if article.number %} ({{ article.number }}),{% endif %}{% if article.start_page and article.end_page %} {{ article.start_page }}-{{ article.end_page }},{% endif %} {% if article.doi %}DOI: {{ article.doi }},{% endif %} {{ article.date|date:"Y" }}.{% endif %}
			{% endfor %}
\end{enumerate}
		{% endif %}
	{% endif %}

{%  else %}
   	No articles found!
{% endif %}

{% endblock %}