### 2016-05-02 templates/report/dissemination/tex/disseminations.tex
NIRA-93 - Generate latex for the dissemination report If the user chooses an internal dissemination, the table has a caption with the chosen internal media outlet.

{% autoescape off %}
{% block content %}

{% if type == 'i' %}
    \begin{longtable}{ |p{4cm}|p{4cm}|p{5cm}|p{2cm}| }
    Title & Topics & URL & Date \\
    \hline
    {% for dissemination in disseminations %}
        {{ dissemination.title }} & {{ dissemination.topics }} & {{ dissemination.link }} & {{ dissemination.date|date:"M d, Y" }} \\
    \hline
    {% endfor %}
    \end{longtable}

{% elif type == 'e' %}
    \begin{longtable}{ |p{3cm}|p{2,3cm}|p{2cm}|p{3cm}|p{3,5cm}|p{1,2cm}| }
    \hline
    \textbf{Title} & \textbf{Authors} & Media outlet & Topics & URL & Date \\
    \hline
    {% for dissemination in disseminations %}
        {{ dissemination.title }} & {{ dissemination.external.authors }} & {{ dissemination.external.media_outlet }} & {{ dissemination.topics }} & {{ dissemination.link }}&#125; & {{ dissemination.date|date:"M d, Y" }} \\
    \hline
    {% endfor %}
    \end{longtable}
{% endif %}

{% endblock %}
{% endautoescape %}