### 2016-06-21 templates/report/dissemination/tex/disseminations.tex
NIRA-105 - Delete the author column of the newsletter report

{% autoescape off %}{% block content %}
{% if type == 'i' and media != "Newsletter" %}\begin{longtable}{ |p{3,5cm}|p{2cm}|p{2,5cm}|p{5cm}|p{2cm}| }
\hline
Title & Authors & Topics & URL & Date \\ \hline
{% for dissemination in disseminations %}{{ dissemination.title }} & {{ dissemination.authors }} & {{ dissemination.topics }} & \url{ {{ dissemination.link }} } & {{ dissemination.date|date:"M d, Y" }} \\ \hline {% endfor %}
\end{longtable}
{% elif type == 'i' and media == "Newsletter" %}\begin{longtable}{ |p{4cm}|p{4cm}|p{5cm}|p{2cm}| }
\hline
Title & Topics & URL & Date \\ \hline
{% for dissemination in disseminations %}{{ dissemination.title }} & {{ dissemination.topics }} & \url{ {{ dissemination.link }} } & {{ dissemination.date|date:"M d, Y" }} \\ \hline {% endfor %}
\end{longtable}
{% elif type == 'e' %}
\begin{longtable}{ |p{3cm}|p{2,3cm}|p{2cm}|p{3cm}|p{3,5cm}|p{1,2cm}| }
\hline
Title & Authors & Media outlet & Topics & URL & Date \\ \hline
{% for dissemination in disseminations %}{{ dissemination.title }} & {{ dissemination.external.authors }} & {{ dissemination.external.media_outlet }} & {{ dissemination.topics }} & \url{ {{ dissemination.link }} } & {{ dissemination.date|date:"M d, Y" }} \\ \hline {% endfor %}
\end{longtable}
{% endif %}{% endblock %}{% endautoescape %}